Individual Variation in Animal Movement and Effects on Seed Dispersal Distance

### 2021-01-08 paper/paper.Rmd
added parameterization paragraph

To understand the consequences of individual variation in animal movement and how these influence seed dispersal distances, we simulated two different scenarios by varying the movement of individual animals in each of those scenarios. We used our first scenario as our null model, in which all the individuals from the simulation had the same average movement rate ($\lambda_1 = \lambda_2 = ... = \lambda_n$), used as the parameter in the probability density function, and thus their movement distances per unit of time were sampled from the same probability distribution ($f(x;\lambda) = \lambda e^{-\lambda x}$). Our second scenario included individual variation in animal movement by incorporating different movement rates for each individual ($\lambda_1 = \lambda_2 = ... = \lambda_n$), which meant that movement distances would be sampled from probability distributions with different parameters ($f(x;\lambda_i) = \lambda_i e^{-\lambda_i x}$). Considering that not all individuals are independent of each other, and that by belonging to social groups their movement patterns might be correlated, we also ran simulations at the scale of family group variation in animal movement by assigning a different movement rate to each social group in the simulations.

@@

Where is the data coming from?
What is a movement rate and how are we calculating it? Why are we using that (uneven time intervals)?
Why are we using an exponential distribution and what type of movement is this? It is an uncorrelated random walk, but the step size is different.

@@

Gut retention time based on data and using a gamma distribution from morales paper.

>>>>

To understand the consequences of individual variation in animal movement and how these influence seed dispersal distances, we simulated three different scenarios by varying the movement of individual animals in each of those scenarios. We used our first scenario as our null model, in which all the individuals from the simulation had the same average movement rate ($\lambda_1 = \lambda_2 = ... = \lambda_n$), used as the parameter in the probability density function, and thus their movement distances per unit of time were sampled from the same probability distribution ($f(x;\lambda) = \lambda e^{-\lambda x}$). Our second scenario included individual variation in animal movement by incorporating different movement rates for each individual ($\lambda_1 = \lambda_2 = ... = \lambda_n$), which meant that movement distances would be sampled from probability distributions with different parameters ($f(x;\lambda_i) = \lambda_i e^{-\lambda_i x}$). Considering that not all individuals are independent of each other, and that by belonging to social groups their movement patterns might be correlated, our third scenario ran simulations at the scale of family group variation in animal movement by assigning a different movement rate to each social group in the simulations.

@@

We focused our study on the many-banded araçari (*Pterglossus pluricinctus*), a small toucan, and its role as a frugivore of the *Virola flexuosa* tree. Previous studies by Holbrook [-@holbrook_home_2011] collected radiotracking information for various dispersers over a period of four years, from 2001 to 2005, in the Ecuadorian Amazon rainforest. The methodology used at the time consisted of capturing and radio-tagging individuals from various toucan species, including *Pterglossus pluricinctus* and two larger Ramphastids, although data on these was scarce. Tracking periods lasted between four and six daylight hours, alternating morning and afternoons, attempting to record bird locations every 15 minutes, a time interval shown to represent the minimum seed retention time for *Virola flexuosa* seeds [@holbrook_using_2007]. Out of the data collected, and following Holbrook [-@holbrook_home_2011], we selected bird individuals with at least 40 recorded point locations. Even though location recordings were attempted every 15 minutes, this was not always possible, due to the individual being too far out from range or canopy cover, therefore point locations data is available every 15 minutes or multiples of 15 minutes, with the majority of point location recordings between 15 and 30 minutes [@holbrook_home_2011]. Further details on field methods can be found in Holbrook [-@holbrook_home_2011].

To take advantage of all the locations recorded, despite this variation in duration of time intervals, we calculated rates of movement, as the average number of meters moved per minute over the entire tracking period for that individual bird. These movement rates were later used in our simulation models as the parameter to describe the probability distribution of movement distances at each one minute time step for each simulation run. In the case of our null model, the movement rate was averaged across all individuals for all the tracking periods, and at the social group level we took the average number of meters moved per minute over the tracking periods for all individuals belonging to that group, with a total of 7 independent groups.

Parameter values for gut retention time were also based on empirically collected data from previous studies [@holbrook_using_2007], were  passage trials were carried out with *Pteroglossus pluricinctus* individuals. However, these passage trials were not performed on every individual captured and some of the trials were also performed with captive birds, thus we could not explore individual variation in gut retention times for the individuals in the movement data set. These trials showed that the average gut retention time for *Pterglossus pluricinctus* was 28 minutes [@holbrook_using_2007], and maximum retention times over 100 minutes. The distribution of gut retention times is characteristically fat-tailed, thus we used a gamma distribution (shape = 4, scale = 5) with the appropriate shift to match our average retention time of 28 minutes [@morales_effects_2006; @levey_effects_2005].

## Model processes

@@
